---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`binomial` is an R package that provides functions for calculating and plotting with binomial random variables.

* `bin_choose()` calculates the number of ways to get k successes out of n trials.
* `bin_probability()` calculates the probability of getting a specified number of successes given the number of trials and probability of success.
* `bin_distribution()` computes the probability distribution given the number of trials and probability of success. Creates a `data.frame` object with additional class `bindis`.
* `bin_cumulative()` is similar to `bin_distribution()`, but also computes cumulative probabilities. Creates a `data.frame` object with additional class `bincum`.
* `bin_variable()` creates a binomial random variable object (class `binvar`) given the number of trials and probability of success.
* `plot()` methods for `bindis` and `bincum` objects
* `summary()` method for `binvar` objects
* `bin_mean()`, `bin_variance()`, `bin_mode()`, `bin_skewness()`, `bin_kurtosis()` compute their respective statistics.

## Motivation

This package was developed as a means to gain practice with creating R packages, writing public/private functions and methods, testing, and documentation.

## Usage

```{r}
library(binomial)

# calculate number of ways to have 2 successes out of 5 trials
bin_choose(n = 5, k = 2)

# 1, 2, or 3 successes out of 5 trials
bin_choose(5, 1:3)

# probability of getting 2 successes in 5 trials, prob of success is 0.5
bin_probability(success = 2, trials = 5, prob = 0.5)

# with 0, 1, and 2 successes
bin_probability(success = 0:2, trials = 5, prob = 0.5)

# compute probability distribution
dis1 <- bin_distribution(trials = 5, prob = 0.5)
dis1

# can use as a regular dataframe or as an object of class "bindis"
class(dis1)

# plot bindis objects
plot(dis1)

# compute cumulative probabilities
dis2 <- bin_cumulative(trials = 5, prob = 0.5)
dis2

class(dis2)

plot(dis2)

# create a binomial random variable
bin1 <- bin_variable(trials = 10, p = 0.3)
bin1

class(bin1)

summary(bin1)

# miscellaneous stats
bin_mean(trials = 10, prob = 0.3)

bin_variance(trials = 10, prob = 0.3)

bin_mode(trials = 10, prob = 0.3)

bin_skewness(trials = 10, prob = 0.3)

bin_kurtosis(trials = 10, prob = 0.3)
```
